Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2025-02-03 Delaneyetal2024.tex
Updates from Overleaf

Greater subglacial friction factors $f_i$ result in greater variability in the shear stress and width-integrate shear stress in the \alpine case (Figure~\ref{fig:params}).
This result is expected given Equations~\ref{eq:tau_g}.
Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
Higher values of $f_i$ allow faster channel growth that accommodates increases in water discharge, reducing velocity variability.
Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).

Smaller values of subaerial channel shape factor $\alpha$, or a  channel cross-section shape closer to a slot canyon, result in greater variability in subaerial velocity and shear stress that can approach the variability of subglacial values (Figure~\ref{fig:params}).
Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}). 
Greater values of subaerial friction factors $f_p$ result in greater variability of shear stress and width-integrated shear stress. 
Yet, they do not exceed the variability of subglaical channels. 
We note that these parameter values span a commonly accepted range.
Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.

@@

The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}). 
Here, we compare the sediment transport behavior of different channel types as they respond to water discharge, channel shape, and hydraulic gradient. 
This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as in subaerial channels.

Channel types examined include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e. R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
However, the formulations establish the scaling of these different channel typesc.
The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.

@@

Several aspects of the models and experiment design make the comparison of the subaerial and subglacial systems difficult. 
The lumped nature of the models means that they operate independently of the upstream drainage network.
Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
However, they neglect more complex, yet important, spatially distributed processes.

>>>>

Greater subglacial friction factors $f_i$ result in greater variability in the shear stress and width-integrate shear stress in the \alpine case (Figure~\ref{fig:params}).
This result is expected given Equations~\ref{eq:tau_g}.
Low values of $f_i$ result in slower growth rates in subglacial channels (Equations~\ref{eq:dS_dt} and \ref{eq:dh}).
As a result, water velocity, as opposed to channel growth, accommodates increases in water discharge, and velocity variability increases.
Higher values of $f_i$ allow faster channel growth that accommodates increase in water discharge, reducing velocity variability.
Smaller values of channel factor $\beta$, creating low and broad channels, result in more variability in width-integrated shear stress.
Here, the channel width can grow more quickly in response to water discharge increases as compared to a semi-circular channel with $\beta = \pi$ (Equation~ \ref{eq:dh}).

Smaller values of subaerial channel shape factor $\alpha$, or a  channel cross-section shape closer to a slot canyon, result in greater variability in subaerial velocity and shear stress that can approach the variability of subglacial values (Figure~\ref{fig:params}).
Steeper subaerial channel slopes $\nabla z_c$ result in greater variability in both shear stress and width-integrated shear stress but only approach the subglacial channel's variability in the \alpine case (Figure~\ref{fig:params}). 
Greater values of subaerial friction factors $f_p$ resulting in greater variability shear stress and width-integrated shear stress. 
Yet, they to not exceed the variability of subglaical channels. 
We note that these parameter values span a commonly accepted range.
Therefore, we do not anticipate a scenario where variability in the subaerial system would exceed the subglacial system with these two hydrologic forcings.

@@

The numerical experiments above consider the size evolution of subglacial channels and demonstrate that for these hydrographs subglacial sediment transport variability is greater than its subaerial counterpart (Section~\ref{sect:ensemble}). 
Here, we compare the sediment transport behavior of different channels types as they respond water discharge, channel shape, and hydraulic gradient. 
This yields insights into the conditions where subglacial sediment transport capacity could exhibit similar variability as subaerial channels.

Channel types examined include subaerial channels (Figure~\ref{fig:chan_types} c), steady-state R-channels \citep[Figure~\ref{fig:chan_types} d]{rothlisberger1972} and pipe-flow (i.e. R-channels of fixed size that do not adjust their size to discharge conditions; Figure~\ref{fig:chan_types} a).
These formulations do not allow us to evaluate non-steady state R-channels presented above (Section~\ref{sect:sub_mode}, Figure~\ref{fig:chan_types} b).
However, the formulations establish the scaling of these different channel types  .
The sediment transport capacity is calculated with a given water discharge and hydraulic gradient for three different sediment transport capacity formulas: Meyer-Peter M\"uller  \citep[MPM; ][]{meyer1948}, Engelund and Hansen \citep[EH; ][]{engelund1967}, and Bagnold \citep{bagnold1980}; additionally width-integrated shear stress is assessed, as used as proxy for sediment transport above.

@@

Several aspects of the models and experiment design make the comparison of the subaerial and subglaical systems difficult. 
The lumped nature of the models means that they operate independently of the upstream drainage network.
Additionally, we omit analysis of suspended sediment discharge records due to their dependence on sediment supply in capturing variations in sediment discharge, in addition to sediment transport capacity \citep[e.g.][]{delaney2019}.
In reality, processes such as sediment access are usually important in controlling sediment export in glacierized catchments \citep[e.g.][]{herman2015,vergara2022}.
Furthermore, meltwater can be distributed to flow through several adjacent conduits impacting the sediment transport capacity in each \citep[e.g.][]{werder2013,hewitt2019,delaney2023}.
The lumped models used here isolate the relationship between water discharge and sediment transport capacity in subglacial and subaerial systems. 
However, they neglect more complex, yet important, spatially distributed processes.
